# ggplot2 {#ggplot2}

## +.gg ()

## aes ()

## aes_ ()

## aes_colour_fill_alpha ()

## aes_group_order ()

## aes_linetype_size_shape ()

## aes_position ()

## annotate ()

## annotation_custom ()

## annotation_logticks ()

## annotation_map ()

## annotation_raster ()

## autoplot ()

## borders ()

## coord_cartesian ()

## coord_fixed ()

## coord_flip ()

## coord_map ()

## coord_polar ()

## coord_trans ()

## cut_interval ()

## diamonds ()

## economics ()

## expand_limits ()

## facet_grid ()

## facet_wrap ()

## faithfuld ()

## fortify ()

## geom_abline ()

## geom_bar ()

## geom_bin2d ()

## geom_blank ()

## geom_boxplot ()

## geom_contour ()

## geom_count ()

## geom_crossbar ()

## geom_density ()

## geom_density_2d ()

## geom_dotplot ()

## geom_errorbarh ()

## geom_freqpoly ()

## geom_hex ()

## geom_jitter ()

## geom_label ()

## geom_map ()

## geom_path ()

## geom_point ()

## geom_polygon ()

## geom_qq ()

## geom_quantile ()

## geom_raster ()

## geom_ribbon ()

## geom_rug ()

## geom_segment ()

## geom_smooth ()

## geom_spoke ()

## geom_violin ()

## ggplot ()

## ggproto ()

## ggsave ()

## ggtheme ()

## guides ()

## guide_colourbar ()

## guide_legend ()

## hmisc ()

## labeller ()

## labellers ()

## label_bquote ()

## labs ()

## lims ()

## luv_colours ()

## margin ()

## mean_se ()

## midwest ()

## mpg ()

## msleep ()

## position_dodge ()

## position_identity ()

## position_jitter ()

## position_jitterdodge ()

## position_nudge ()

## position_stack ()

## presidential ()

## print.ggplot ()

## print.ggproto ()

## qplot ()

## resolution ()

## scale_alpha ()

## scale_colour_brewer ()

## scale_colour_gradient ()

## scale_colour_grey ()

## scale_colour_hue ()

## scale_continuous ()

## scale_date ()

## scale_identity ()

## scale_linetype ()

## scale_manual ()

Read in a CSV file with survey data about Thanksgiving. 

More info available here: 

https://github.com/fivethirtyeight/data/tree/master/thanksgiving-2015

```{r}
csv_link <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/thanksgiving-2015/thanksgiving-2015-poll-data.csv"
thanksgiving <- read.csv(csv_link,header=TRUE,stringsAsFactors=FALSE)
```

Filter only for individuals that celebrate Thanksgiving, after cleaning up some column names.

```{r}
colnames(thanksgiving)[2] <- "Celebrates_Thanksgiving"
colnames(thanksgiving)[c(25,38,50)] <- c("Which of these side dishes aretypically served at your Thanksgiving dinner? Please select all that apply. - Other (please specify again)",
"Which type of pie is typically served at your Thanksgiving dinner? Please select all that apply. - Other (please specify again)",
"Which of these desserts do you typically have at Thanksgiving dinner? Please select all that apply.   - Other (please specify again)")
thanksgiving <- thanksgiving %>% filter(Celebrates_Thanksgiving == "Yes")
```

Select columns corresponding to side dishes, minus "Other".

Then, count number of people who selected each side.

```{r}
side_dish_columns <- grep('Which of these side dishes aretypically served at your Thanksgiving dinner?',colnames(thanksgiving),value=TRUE)
side_dish_columns <- side_dish_columns[grep('Other',side_dish_columns,invert=TRUE)]

side_dish_frequency <- data.frame(Side = rep(NA,times=length(side_dish_columns)),
Frequency = rep(NA,times=length(side_dish_columns)))

for(i in 1:length(side_dish_columns))
{
side_name <- str_replace_all(side_dish_columns[i],
pattern='Which of these side dishes aretypically served at your Thanksgiving dinner\\? Please select all that apply\\. - ',
replace='')

side_dish_frequency[i,] <- c(side_name,length(which(thanksgiving[,side_dish_columns[i]] != "")))
}

side_dish_frequency$Frequency <- as.numeric(side_dish_frequency$Frequency)
```

Make a barplot of the frequency of different sides, minus "Other".

```{r}
ggplot(side_dish_frequency,aes(Side,Frequency)) + 
geom_bar(stat="identity") + 
theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Now use scale_fill_manual to color each bar by a color appropriate to the food.

Use colors here as a guide: http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

Put in colors based on order sides appear in side_dish_frequency.

Let's color fruit salad as red, since may have watermelon, strawberry, etc.

```{r}
ggplot(side_dish_frequency,aes(Side,Frequency,fill=Side)) + 
geom_bar(stat="identity") + 
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
scale_fill_manual(values = c("darkgreen","darkorange","floralwhite","gold","gold","firebrick1","darkgreen","gold","floralwhite","floralwhite","darkorange","darkgreen","darkorange"))
```

Sort by frequency, then adjust the order of the color vector accordingly.

```{r}
color_per_side <- c("darkgreen","darkorange","floralwhite","gold","gold","firebrick1","darkgreen","gold","floralwhite","floralwhite","darkorange","darkgreen","darkorange")
color_per_side <- color_per_side[order(side_dish_frequency$Frequency)]

side_dish_frequency$Side <- factor(side_dish_frequency$Side,
levels=side_dish_frequency$Side[order(side_dish_frequency$Frequency)])

ggplot(side_dish_frequency,
aes(Side,Frequency,fill=Side)) + 
geom_bar(stat="identity") + 
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
scale_fill_manual(values = color_per_side)
```

We find that the two most popular Thanksgiving sides are both white.

What if we use shape_colour_manual instead, and specify color by side instead of fill by side?

```{r}
ggplot(side_dish_frequency,
aes(Side,Frequency,colour=Side)) +
geom_bar(stat="identity",fill="lightgrey") + 
theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
scale_colour_manual(values = color_per_side)
```

Now the color of the border of the bars will change based on the colors we have specified.

Using "colour" instead of "fill" probably isn't what we want for a barplot. But it might be useful for other types of plots (like scatterplots with a hollow dot for the points).

## scale_shape ()

## scale_size ()

## scale_x_discrete ()

## seals ()

## sec_axis ()

## stat_ecdf ()

## stat_ellipse ()

## stat_function ()

## stat_identity ()

## stat_summary_2d ()

## stat_summary_bin ()

## stat_unique ()

## theme ()

## theme_get ()

## txhousing ()
