# dplyr {#dplyr}

dplyr provides a grammar of data manipulation, providing a consistent set of verbs that solve the most common data manipulation challenges. 

## select()

The select function provides options for selecting only certain columns of data to work with.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
alcohol.data <- read_csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/alcohol-consumption/drinks.csv') 
alcohol.data %>% head() %>% kable()
```

At it's most basic, select allows for the manual selection of columns

```{r}
alcohol.data %>%
  select(country, total_litres_of_pure_alcohol) %>%
  head() %>%
  kable()
```

```{r}
alcohol.data %>%
  select(c(1, 3)) %>%
  head() %>%
  kable()
```

```{r}
alcohol.data %>%
  select(1:4) %>%
  head() %>%
  kable()
```

There are a number of additional functions that can be used to select columns based on specific criteria. These include:

* ```starts_with()``` which selects all columns that begin with the given text
* ```ends_with()``` which selects all columns that end with the given text
* ```contains()``` which selects all columns that contain the given text anywhere
* ```matches()``` which selects all columns that match the given regular expression. See regular expressions for more details.
* ```num_range("x", 1:5)``` which selects the variables named x01, x02, x03, x04 and x05
* ```one_of(x)``` which selects every name that appears in x, which should be a character vector.

Finally, a dash (-) may be added to the beginning of any of these functions to select all columns that do NOT match.

```{r}
alcohol.data %>%
  select(-ends_with('servings')) %>%
  head() %>%
  kable()
```

The select function can also be used to rearrange the columns in a data frame. 

```{r}
alcohol.data %>%
  select(1, 5, 2:4) %>%
  head() %>%
  kable()
```

A special function ```everything()``` will select all remaining columns in order. This is best used when you wish to move a single column to a new position but do not care to rearrange the remaining columns.

```{r}
alcohol.data %>%
  select(1, 5, everything()) %>%
  head() %>%
  kable()
```

Select also has the ability to rename columns, however, it will drop all columns not mentioned. In order to rename columns but to keep all columns (regardless of whether their names have been changed) it is best to use ```rename()```. See rename().

```{r}
alcohol.data %>%
  select(location=country, everything()) %>%
  head() %>%
  kable()
```

## rename()

The rename function is used to alter the names of columns. Any columns not specified will not be changed. If you wish to drop unnamed columns at the same time, use ```select()``` See select().

```{r}
alcohol.data %>%
  rename(beer=beer_servings, spirit=spirit_servings, wine=wine_servings) %>% 
  head() %>%
  kable()
```

## filter()

The filter function allows a user to subset a data frame based on the values in a given column.

```{r, warning=FALSE, message=FALSE}
avengers.data <- read_csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv') %>%
  select(2:6)
avengers.data %>% head() %>% kable()
```

Perhaps the user would like to work only with the female avengers.

```{r}
avengers.data %>%
  filter(Gender == 'FEMALE') %>%
  head() %>%
  kable()
```

Perhaps the user would like to work only with female avengers with more than 500 appearances.

```{r}
avengers.data %>%
  filter(Gender == 'FEMALE' & Appearances > 500) %>%
  head() %>%
  kable()
```

There are numerous comparison options that can be used with filter. These include: ```==, >, >=, <, <=, !=```
The use of a single ampersand ```&``` is used to indicate 'and' while a single pipe ```|``` is used to indicate 'or'. 
The exclamation point ```!``` is used to indicate 'nor' (see next example) and ```xor()``` is the exclusive or.

Filtering can also be based on whether an element is NA.

This data set does not have any NAs, but rather has empty strings in cases where the variable does not apply. 

To demonstrate the technique of filtering out NAs, let's convert these empty strings to NAs.

Then, use the filter function to select cases where the variable is non-NA.

```{r}
avengers.data_incl_NAs <- data.frame(avengers.data,check.names=FALSE)
var <- "Probationary Introl"
avengers.data_incl_NAs[,var] <- plyr::mapvalues(avengers.data_incl_NAs[,var],
  from = "",to=NA)

avengers.data_incl_NAs %>%
filter(!is.na(`Probationary Introl`)) %>%
kable()
```

Between can be used to filter between a range of values.

```{r}
avengers.data %>%
  filter(between(Appearances, 500, 1000)) %>%
  head() %>%
  kable()
```

Finally, near can be used to select a range of values within a certain range of a central value.

```{r}
avengers.data %>%
  filter(near(Appearances, 612, tol=20)) %>%
  head() %>%
  kable()
```

Of course, all of these can be combined with 'and' ```&``` and 'or' ```|``` logical operators.

There are three related functions ```filter_if()```, ```filter_all()```, and ```filter_at()```. See scoped operations.

For a last example, let's combine the and operator with the or operator to choose results meeting two mutually exclusive criteria, each based on two different conditions.

```{r}
high_appearance_females <- avengers.data %>%
filter(Gender == "FEMALE" & Appearances > 500)

low_appearance_males <- avengers.data %>%
filter(Gender == "MALE" & Appearances < 200)

high_appearance_females_or_low_appearance_males <- avengers.data %>%
filter((Gender == "FEMALE" & Appearances > 500) | (Gender == "MALE" & Appearances < 200))

nrow(high_appearance_females)
nrow(low_appearance_males)
nrow(high_appearance_females_or_low_appearance_males)
```

There are 10 high appearance females and 65 low appearance males. Thus, the fact that we get 75 results for the query using OR operator makes sense, as the criteria we looked for using the OR operator are mutually exclusive.

## arrange()

arrange changes the natural column order of the data frame. It takes data frame
name and the column name(s) as input and re arranges the columns based on input.


## mutate() and transmute()

The mutate function creates new columns which can be added to a copy of the dataset.  You can also add new columns to your existing dataset by overwriting it like this:

```{r}
alcohol.data <- mutate(alcohol.data, total_servings = beer_servings + spirit_servings + wine_servings)
```

You can use any mathematical operator (+, -, *, /) as well as functions like:

- log()
- lead(), lag()
- dense_rank(), min_rank(), percent_rank(), row_number(), cume_dist(), ntile()
- cumsum(), cummean(), cummin(), cummax(), cumany(), cumall()
- na_if(), coalesce()
- if_else(), recode(), case_when()

```{r}
# Rank into 5 "buckets" by displacement
mtcars <- mutate(mtcars, rank = ntile(disp, 5))
```

Transmute does the same thing as mutate, but drops all the other columns not specified in your function arguments.

```{r}
transmute(alcohol.data, country, total_servings = beer_servings + spirit_servings + wine_servings)
```

## summarise()

The summarise function returns a table of summary statistics that you specify by runnning functions on your data.  You name each statistic and then specify what function to use and separate additional statistics with commas like this:

```{r}
summarise(mtcars, avg_disp = mean(disp), max_disp = max(disp), min_disp = min(disp), median_disp = median(disp))
```

In addtion to all the usual summary statistic functions in R, summarise comes with some addtional functions that you can use:

* ```first()```  - returns the first value
* ```last``` - returns the last value
* ```nth``` - returns the nth value
* ```n``` - returns the number of rows
* ```n_distinct``` - returns the number of unique values

```{r}
summarise(mtcars, first = first(disp), last = last(disp), fifth = nth(disp, 5), n = n(), unique = n_distinct(disp))
```

## sample_n() and sample_frac()

`sample_n` and `sample_frac` return random samples from your data.  `sample_n` wil select the number of rows you specify in the second argument at random. `sample_frac` returns a random sample fraction of rows.  You specify the fraction of rows you want returned in your second function argument with a decimal number from 0 to 1.  Both functions can be set to sample with or without replacement using the `replace = TRUE/FALSE` argument.  

```{r}
sample_n(mtcars, 5, replace = TRUE)
```

```{r}
sample_frac(mtcars, 0.25, replace = FALSE)
```

