# tidyr {#tidyr}

## complete ()

```{r}
library(dplyr)
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
```

Complete a data frame with missing combinations of data.

Description:

Turns implicit missing values into explicit missing values. This is a wrapper around expand(), dplyr::left_join() and replace_na() that’s useful for completing missing combinations of data.

Usage:

    complete(data, ..., fill = list())

Example:

```{r}
df <- tibble(
  group = c(1:2, 1),
  item_id = c(1:2, 2),
  item_name = c("a", "b", "b"),
  value1 = 1:3,
  value2 = 4:6
)

df %>% complete(group, nesting(item_id, item_name))

# You can also choose to fill in missing values
df %>% complete(group, nesting(item_id, item_name), fill = list(value1 = 0))
```

## drop_na ()

Drop rows containing missing values

Description:

Drop rows containing missing values

Usage:

    drop_na(data, ...)

Example:

```{r}
df <- tibble(x = c(1, 2, NA), y = c("a", NA, "b"))
df %>% drop_na()
df %>% drop_na(x)
```

## expand ()

Expand data frame to include all combinations of values

Description:

expand() is often useful in conjunction with left_join if you want to convert implicit missing values to explicit missing values. Or you can use it in conjunction with anti_join() to figure out which combinations are missing.

Usage:

    expand(data, ...)

Example:

```{r}
# All possible combinations of vs & cyl, even those that aren't
# present in the data
expand(mtcars, vs, cyl)
```

## extract ()

Extract one column into multiple columns.

Description:

Given a regular expression with capturing groups, extract() turns each group into a new column. If the groups don’t match, or the input is NA, the output will be NA.

Usage:

    extract(data, col, into, regex = "([[:alnum:]]+)", remove = TRUE,
      convert = FALSE, ...)

Example:

```{r}
df <- data.frame(x = c(NA, "a-b", "a-d", "b-c", "d-e"))
df %>% extract(x, "A")
df %>% extract(x, c("A", "B"), "([[:alnum:]]+)-([[:alnum:]]+)")
```

## fill ()

Fill in missing values.

Description:

Fills missing values in using the previous entry. This is useful in the common output format where values are not repeated, they’re recorded each time they change.

Usage:

     fill(data, ..., .direction = c("down", "up"))

Example:

```{r}
df <- data.frame(Month = 1:12, Year = c(2000, rep(NA, 11)))
df %>% fill(Year)  
```

## full_seq ()

## gather ()

Gather columns into key-value pairs.

Description:

Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed. You use gather() when you notice that you have columns that are not variables.

Usage:

    gather(data, key = "key", value = "value", ..., na.rm = FALSE,
      convert = FALSE, factor_key = FALSE)

Example:

```{r}
# From http://stackoverflow.com/questions/1181060
stocks <- tibble(
  time = as.Date('2009-01-01') + 0:9,
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4))
gather(stocks, stock, price, -time)
stocks %>% gather(stock, price, -time)
```

## nest ()

Nest repeated values in a list-variable.

Description:

There are many possible ways one could choose to nest columns inside a data frame. nest() creates a list of data frames containing all the nested variables: this seems to be the most useful form in practice.

Usage:

     nest(data, ..., .key = "data")

Example:

```{r}
as_tibble(iris) %>% nest(-Species)
as_tibble(chickwts) %>% nest(weight)
if (require("gapminder")) {
  gapminder %>%
    group_by(country, continent) %>%
    nest()
  gapminder %>%
    nest(-country, -continent)
}     
```

## replace_na ()

Replace missing values

Description:

Replace missing values

Usage:

     replace_na(data, replace, ...)

Example:

```{r}
df <- tibble(x = c(1, 2, NA), y = c("a", NA, "b"), z = list(1:5, NULL, 10:20))
df %>% replace_na(list(x = 0, y = "unknown"))

# NULL are the list-col equivalent of NAs
df %>% replace_na(list(z = list(5)))
df$x %>% replace_na(0)
df$y %>% replace_na("unknown")
```

## separate ()

Separate one column into multiple columns.

Description:

Given either regular expression or a vector of character positions, separate() turns a single char- acter column into multiple columns.

Usage:

     separate(data, col, into, sep = "[^[:alnum:]]+", remove = TRUE,
      convert = FALSE, extra = "warn", fill = "warn", ...)

Example:

```{r}
df <- data.frame(x = c(NA, "a.b", "a.d", "b.c"))
df %>% separate(x, c("A", "B"))

# If every row doesn't split into the same number of pieces, use
# the extra and file arguments to control what happens
df <- data.frame(x = c("a", "a b", "a b c", NA))
df %>% separate(x, c("a", "b"))

# The same behaviour but no warnings
df %>% separate(x, c("a", "b"), extra = "drop", fill = "right")

# Another option:
df %>% separate(x, c("a", "b"), extra = "merge", fill = "left")

# If only want to split specified number of times use extra = "merge"
df <- data.frame(x = c("x: 123", "y: error: 7"))
df %>% separate(x, c("key", "value"), ": ", extra = "merge")
```

## separate_rows ()

## smiths ()

## spread ()

Take a tidy key-value pairs and spread them across multiple columns.

Description:

The antithesis of gather, spread() unstacks or unpivots tables by spreading rows into columns.

Usage:

  spread(data, key, value, fill = NA, convert = FALSE, drop = TRUE, sep = NULL)

Example:

```{r}
alcohol.data <- read_csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/alcohol-consumption/drinks.csv')

alcohol.data %>% head() %>% kable()

# While gather() "pivots" the data table so that each observation forms a row...
alc.gather <- alcohol.data %>% gather("type", "amount", 2:5)
alc.gather %>% head() %>% kable()                        

# ...spread() performs the opposite operation, widening the data table.
alc.spread <- alc.gather %>% spread(type, amount)
alc.spread %>% head() %>% kable()
```
## table1 ()

## uncount ()

## unite ()


Description: joins multiple columns together by concatenation and seperates values by a seperator


Usage:unite(data, col,..., sep = "_", remove = TRUE)


data = Dataframe

col =  Desired name of new column can be an unquoted string

... =  A list of columns to Unite

sep =  How to seperate the united vlaues

Example:  Use unite function to create Member index in Congress_age DF.
Member will be a concatination of firstname and last name columns with a " " as a seperator

```{r}
Congress_Age <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/congress-age/congress-terms.csv")
  Congress_Age <- Congress_Age %>%
      unite(.,Member,c(firstname,lastname),sep=" ")

## print for confirmation
  kable(head(Congress_Age$Member),caption="New Member column")
```
<br>

## unnest ()

**Description:** If you have a column with lists of items (a list-column), `unnest()` makes each list element its own row.

**Usage:** `unnest (data, ..., .drop = NA, .id = NULL, .sep = NULL, .preserve = NULL)`

* `data` --- a data frame.

* `...` --- the columns to unnest; defaults to all list-columns

* `.drop` --- whether additional list columns should be dropped

* `.id` --- data frame identifier; creates a new column with the name `.id`, giving a unique identifier.

* `.sep` --- identify a separator to use in the names of unnested data frame columns, which combine the name of the original list-col with the names from nest data frame

* `.preserve` --- list-columns to preserve in the output

```{r}

# Read in the test dataframe and filter for movies with multiple directors listed
directors <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/biopics/biopics.csv") %>% 
             filter(str_detect(director, ".\\,.")) %>% 
             select(title, country, director)

# Examine the nested dataframe
head(directors, 3) %>% 
  kable("html")  %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# Apply the unnest function, and identify on which character to split the strings in the column-list
dir_unnest <- unnest(directors, director = strsplit(director, ","))

# Examine the unnested dataframe
head(dir_unnest, 7) %>% 
  kable("html")  %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```

<br>

## who ()
