# readr {#readr}

```{r}
library(tidyverse)
library(hms)
```
## read_csv()

    ** read_csv() - Reads comma-delimited files

    ** read_csv2() - Reads semicolon-separated files

    ** read_tsv() - Reads tab-delimited files

    ** read_delim() - Reads in files with any delimiter

Read file will print out a column specification that gives the name and type of each column:

```{r}
data <- read_csv("data/example.csv")
```

`read_csv()` uses the first line of the data for the column names, the following examples will show how to tweak this behavior:

    * Using skip = n (n is the number of lines to skip)

    ```{r}
    read_csv("This the first line
    This is the second line
    a,b,c
    1,2,3", skip = 2)
    ```

    * Using comment = "#" to skip all lines that starts with "#"

    ```{r}
    read_csv("# This is a comment line
    a,b,c
    1,2,3", comment = "#")
    ```

    * If there's no column names, using col_names = FALSE

    ```{r}
    read_csv("1,2,3
    4,5,6", col_names = FALSE)
    ```

    * If there's column names

    ```{r}
    read_csv("1,2,3
    4,5,6", col_names = c("a", "b", "c"))
    ```

## cols ()

## cols_condense ()

## col_skip ()

## count_fields ()

## date_names ()

## format_delim ()

## guess_encoding ()

To find the types of encoding:

```{r}
guess_encoding(charToRaw("Data Analytics"))
```

## locale ()

The goal of readr’s locales is to encapsulate common options that vary between languages and localities. This includes:

    - The names of months and days, used when parsing dates.
    - The default time zone, used when parsing datetimes.
    - The character encoding, used when reading non-ASCII strings.
    - Default date format, used when guessing column types.
    - The decimal and grouping marks, used when reading numbers.

## parse_atomic ()

## parse_datetime (), parse_date() and parse_time()

Allow to parse varoious date and time specifications.

```{r}
str(parse_date(c("2018-05-01", "2018-04-30")))
```

```{r}
parse_time(c("6:43 pm"))
```

## parse_factor ()

Creates factors, the data structure that R uses to represent categorical variables with fixed and known values.

```{r}
parse_factor(c("a", "b", "c"), levels = c("a", "b"))
```

## parse_logical(), parse_number(), parse_double() and parse_integer()

Parse logicals, numbers, doubles and integers respectively.

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA")))
```

```{r}
parse_number("$112,33")
parse_number("$12.33.21", locale = locale(grouping_mark = "."))
```

```{r}
parse_double("12.33")
parse_double("12,33", locale = locale(decimal_mark = ","))
```

```{r}
str(parse_integer(c("1", "2", "3")))
```

## parse_character()

```{r}
parse_character("Data analytics", locale = locale(encoding = "Latin1"))
```

## parse_guess ()

## problems ()

## read_delim ()

## read_file ()

## read_fwf ()

## read_lines ()

## read_log ()

## read_table ()

## spec_delim ()

## type_convert ()

## write_delim ()
