# purrr {#purrr}

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)

libs <- c("tidyverse", "magrittr", "knitr", "kableExtra")
lapply(libs, require, character.only = T)
```

```{r dataload, include=FALSE}
cars <- mtcars
mpg <- cars[["mpg"]]
```



## Overview
purrr enhances R’s functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors. If you’ve never heard of FP before, the best place to start is the family of map() functions which allow you to replace many for loops with code that is both more succinct and easier to read.

## Map Family

The complete list of functions within the map family are:
```
map(.x, .f, ...)

map_lgl(.x, .f, ...)

map_chr(.x, .f, ...)

map_int(.x, .f, ...)

map_dbl(.x, .f, ...)

map_dfr(.x, .f, ..., .id = NULL)

map_dfc(.x, .f, ...)

walk(.x, .f, ...)
```

Arguments

- .x: A list or atomic vector
- .f: A function, formula, or atomic vector
    - If .f is a function, it is used as is
    - If .f is a formula, e.g. `~ .x + 2`, it is converted to a function. There are three ways to refer to the arguments:
        - For a single argument function, use .
        - For a two argument function, use .x and .y
        - For more arguments, use ..1, ..2, ..3, etc...


### map

The map function transforms the input, returning a vector of the same length as the input. map, along with reduce, appear to draw their inspiration from JavaScript. In R, map can be very useful when the programmers wants to use the same function on a list. dplyr has mutate for working with data frames, while purrr specializes in lists.

#### Function Example
Lets extract a vector of numbers from an mtcars dataset (mpg). Lets use map to convert that vector into kmpg
```{r}
kmpg <- mpg %>%
  map(.f = function(x) x * 1.60934) %>%
  print


data_frame(mpg = c(class(mpg)), kmpg = c(class(kmpg))) %>%
  kable
```

Using map in this instance, returned a list. This may not be preferable in some situations and purrr has built in functions that can return exactly what you are expecting. Our original vector was a numeric vector. We can return a numeric vector using the map_dbl function. Lets re-try the code above using map_dbl.

```{r}
kmpg <- mpg %>%
  map_dbl(.f = function(x) x * 1.60934) %>%
  print


data_frame(mpg = c(class(mpg)), kmpg = c(class(kmpg))) %>%
  kable
```

As shown, both mpg and kmpg are both numerical vectors. purrr also comes supplies with map_chr, map_dbl, map_dfc (column bind), map_dfr (row bind), map_int, map_lgl. All return vectors except map_dfc and map_dfr which return data frames.

#### Formula Example

As stated above, we can also use the map functions with the formula syntax
```{r}
mpg %>%
  map_dbl(.f = ~ .x * 1.60934) %>%
  print
```

## Reduce Family

The complete list of functions within the reduce family are:
```
reduce(.x, .f, ..., .init)

reduce_right(.x, .f, ..., .init)

reduce2(.x, .y, .f, ..., .init)

reduce2_right(.x, .y, .f, ..., .init)
```

Arguments

- .x: A list or atomic vector
- .f: A 2-argument function. The function will be passed the accumulated value as the first argument and the next value as the second argument. For reduce2(), a 3-argument function. The function will be passed the accumulated value as the first argument, the next value of .x as the second argument, and the next value of .y as the third argument.
- ...: Additional arguments passed on to .f
- .init: If supplied, will be used as the first value to start the accumulation, rather than using x[[1]]. This is useful if you want to ensure that reduce returns a correct value when .x is empty. If missing, and x is empty, will throw an error
- .y: For reduce2(), an additional argument that is passed to .f. If init is not set, .y should be 1 element shorter than .x

### reduce

The reduce function transforms the input into a single value by iteratively applying a binary function. Another function whose inspiration seems to stem from JavaScript. In R, reduce can be useful when you have a list or vector of values that need to be 
"accumulated" in some form to create a single item.

#### Example
Lets pretend that we wanted to create a pipe-delimited string from the row names of our cars data
```{r}
names <- row.names(cars) %>%
  reduce(.f = function(start, .next) paste(start, .next, sep = "|")) %>%
  print
```

