# stringr {#stringr}
```{r setup}
library(stringr)
```

###Pattern matching

####str_count
Count the number of matches in a string

Examples
```{r}
fruit <- c("apple", "banana", "pear", "pineapple")
str_count(fruit, "a")
str_count(fruit, "p")
str_count(fruit, "e")
str_count(fruit, c("a", "b", "p", "p"))
str_count(c("a.", "...", ".a.a"), ".")
str_count(c("a.", "...", ".a.a"), fixed("."))
```

####str_detect
Detect the presence or absence of a pattern in a string.

Examples
```{r str_detect}
fruit <- c("apple", "banana", "pear", "pinapple")
str_detect(fruit, "a")
str_detect(fruit, "^a")
str_detect(fruit, "a$")
str_detect(fruit, "b")
str_detect(fruit, "[aeiou]")
# Also vectorised over pattern
str_detect("aecfg", letters)
```

####str_extract
Extract matching patterns from a string.

Examples:
```{r string extract}
shopping_list <- c("apples x4", "bag of flour", "bag of sugar", "milk x2")
str_extract(shopping_list, "\\d")

str_extract(shopping_list, "[a-z]+")

str_extract(shopping_list, "[a-z]{1,4}")

str_extract(shopping_list, "\\b[a-z]{1,4}\\b")

# Extract all matches
str_extract_all(shopping_list, "[a-z]+")

str_extract_all(shopping_list, "\\b[a-z]+\\b")

str_extract_all(shopping_list, "\\d")

# Simplify results into character matrix
str_extract_all(shopping_list, "\\b[a-z]+\\b", simplify = TRUE)

str_extract_all(shopping_list, "\\d", simplify = TRUE)

# Extract all words
str_extract_all("This is, suprisingly, a sentence.", boundary("word"))
```

####str_extract_all
str_extract_all(string, pattern, simplify = FALSE)
Examples:

```{r str_extract-all}
shopping_list <- c("apples x4", "bag of flour", "bag of sugar", "milk x2")

str_extract(shopping_list, "\\d")

str_extract(shopping_list, "[a-z]+")

str_extract(shopping_list, "[a-z]{1,4}")

str_extract(shopping_list, "\\b[a-z]{1,4}\\b")

# Extract all matches
str_extract_all(shopping_list, "[a-z]+")

str_extract_all(shopping_list, "\\b[a-z]+\\b")

str_extract_all(shopping_list, "\\d")

str_extract_all(shopping_list, "\\b[a-z]+\\b", simplify = TRUE)

str_extract_all(shopping_list, "\\d", simplify = TRUE)

str_extract_all("This is, suprisingly, a sentence.", boundary("word"))

```

###Locate the position of patterns in a string.
####str_locate(string, pattern)
For str_locate, an integer matrix. First column gives start postion of match, and second column gives end position. For str_locate_all a list of integer matrices.

Vectorised over string and pattern. If the match is of length 0, (e.g. from a special match like $) end will be one character less than start.

Examples str_locate and str_locate_all:
```{r str_locate}
fruit <- c("apple", "banana", "pear", "pineapple")
str_locate(fruit, "$")

str_locate(fruit, "a")

str_locate(fruit, "e")

str_locate(fruit, c("a", "b", "p", "p"))

str_locate_all(fruit, "a")

str_locate_all(fruit, "e")

# Find location of every character
str_locate_all(fruit, "")

```

###Extract matched groups from a string.

Vectorised over string and pattern.
str_match(string, pattern)
str_match_all(string, pattern)

For str_match, a character matrix. First column is the complete match, followed by one column for each capture group. For str_match_all, a list of character matrices.

Examples:
```{r str_match}


strings <- c(" 219 733 8965", "329-293-8753 ", "banana", "595 794 7569",
  "387 287 6718", "apple", "233.398.9187  ", "482 952 3315",
  "239 923 8115 and 842 566 4692", "Work: 579-499-7527", "$1000",
  "Home: 543.355.3679")
phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"

str_extract(strings, phone)

str_match(strings, phone)

# Extract/match all
str_extract_all(strings, phone)

str_match_all(strings, phone)
x <- c("<a> <b>", "<a> <>", "<a>", "", NA)
str_match(x, "<(.*?)> <(.*?)>")
str_match_all(x, "<(.*?)>")

str_extract(x, "<.*?>")

str_extract_all(x, "<.*?>")


```
### str_to_upper ()
str_to_upper (string)
Convert case of a string.

Example:
```{r str_to_upper} 
exstring <- "the example string"
str_to_upper(exstring)
```

## invert_match ()
invert_match (string)

Invert a matrix of match locations to match the opposite of what was previously matched.

Example:
```{r invert_match} 
numbers <- "1 and 2 and 4 and 456"
num_loc <- str_locate_all(numbers, "[0-9]+")[[1]]
str_sub(numbers, num_loc[, "start"], num_loc[, "end"])

text_loc <- invert_match(num_loc)
str_sub(numbers, text_loc[, "start"], text_loc[, "end"])

```

## modifiers ()

## stringr-data ()

## str_c ()

## str_conv ()
str_conv(string, encoding)

This is a convenient way to override the current encoding of a string.

```{r str_conv}
x <- rawToChar(as.raw(177))
x
str_conv(x, "ISO-8859-2")
str_conv(x, "ISO-8859-1")
```

## str_count ()

str_count(string, pattern = "")

Count The Number Of Matches In A String.

Example:-

```{r str_count}
fruit <- c("apple", "banana", "pear", "pineapple")
str_count(fruit, "a")
str_count(fruit, "p")
str_count(fruit, "e")
str_count(fruit, c("a", "b", "p", "p"))
```


## str_dup ()

str_dup(string, times)

Duplicate And Concatenate Strings Within A Character Vector.

Example:-

```{r str_dup}
fruit <- c("apple", "pear", "banana")
str_dup(fruit, 2)
str_dup(fruit, 1:3)
str_c("ba", str_dup("na", 0:5))
```


## str_flatten ()

str_flatten(string, collapse = "")

Flatten a string

Example:-

```{r str_flatten}
str_flatten(letters)
str_flatten(letters, "-")
```


## str_glue ()

str_glue(..., .sep = "", .envir = parent.frame())

These functions are wrappers around glue::glue() and glue::glue_data(), which provide a powerful and elegant syntax for interpolating strings. These wrappers provide a small set of the full options. Use the functions directly from glue for more control.

Example:-

```{r str_glue}
name <- "Fred"
age <- 50
anniversary <- as.Date("1991-10-12")
str_glue(
  "My name is {name}, ",
  "my age next year is {age + 1}, ",
  "and my anniversary is {format(anniversary, '%A, %B %d, %Y')}."
)
```
## str_length ()

str_length(string)

Returns The Length Of A String.

Example:-

```{r str_length}
str_length(c("i", "like", "Ice Cream", NA))
```


## str_locate ()

## str_match ()

## str_order ()

## str_pad ()

## str_remove ()

## str_replace ()

## str_replace_na ()

## str_split ()

## str_sub ()

## str_subset ()

## str_trim ()

## str_trunc ()

## str_view ()

## str_wrap ()

## word ()
