# stringr {#stringr}
```{r setup}
library(stringr)
```

###Pattern matching

####str_count
Count the number of matches in a string

Examples
```{r}
fruit <- c("apple", "banana", "pear", "pineapple")
str_count(fruit, "a")
str_count(fruit, "p")
str_count(fruit, "e")
str_count(fruit, c("a", "b", "p", "p"))
str_count(c("a.", "...", ".a.a"), ".")
str_count(c("a.", "...", ".a.a"), fixed("."))
```

####str_detect
Detect the presence or absence of a pattern in a string.

Examples
```{r str_detect}
fruit <- c("apple", "banana", "pear", "pinapple")
str_detect(fruit, "a")
str_detect(fruit, "^a")
str_detect(fruit, "a$")
str_detect(fruit, "b")
str_detect(fruit, "[aeiou]")
# Also vectorised over pattern
str_detect("aecfg", letters)
```

####str_extract
Extract matching patterns from a string.

Examples:
```{r string extract}
shopping_list <- c("apples x4", "bag of flour", "bag of sugar", "milk x2")
str_extract(shopping_list, "\\d")

str_extract(shopping_list, "[a-z]+")

str_extract(shopping_list, "[a-z]{1,4}")

str_extract(shopping_list, "\\b[a-z]{1,4}\\b")

# Extract all matches
str_extract_all(shopping_list, "[a-z]+")

str_extract_all(shopping_list, "\\b[a-z]+\\b")

str_extract_all(shopping_list, "\\d")

# Simplify results into character matrix
str_extract_all(shopping_list, "\\b[a-z]+\\b", simplify = TRUE)

str_extract_all(shopping_list, "\\d", simplify = TRUE)

# Extract all words
str_extract_all("This is, suprisingly, a sentence.", boundary("word"))
```

####str_extract_all
str_extract_all(string, pattern, simplify = FALSE)
Examples:

```{r str_extract-all}
shopping_list <- c("apples x4", "bag of flour", "bag of sugar", "milk x2")

str_extract(shopping_list, "\\d")

str_extract(shopping_list, "[a-z]+")

str_extract(shopping_list, "[a-z]{1,4}")

str_extract(shopping_list, "\\b[a-z]{1,4}\\b")

# Extract all matches
str_extract_all(shopping_list, "[a-z]+")

str_extract_all(shopping_list, "\\b[a-z]+\\b")

str_extract_all(shopping_list, "\\d")

str_extract_all(shopping_list, "\\b[a-z]+\\b", simplify = TRUE)

str_extract_all(shopping_list, "\\d", simplify = TRUE)

str_extract_all("This is, suprisingly, a sentence.", boundary("word"))

```

###Locate the position of patterns in a string.
####str_locate(string, pattern)
For str_locate, an integer matrix. First column gives start postion of match, and second column gives end position. For str_locate_all a list of integer matrices.

Vectorised over string and pattern. If the match is of length 0, (e.g. from a special match like $) end will be one character less than start.

Examples str_locate and str_locate_all:
```{r str_locate}
fruit <- c("apple", "banana", "pear", "pineapple")
str_locate(fruit, "$")

str_locate(fruit, "a")

str_locate(fruit, "e")

str_locate(fruit, c("a", "b", "p", "p"))

str_locate_all(fruit, "a")

str_locate_all(fruit, "e")

# Find location of every character
str_locate_all(fruit, "")

```

###Extract matched groups from a string.

Vectorised over string and pattern.
str_match(string, pattern)
str_match_all(string, pattern)

For str_match, a character matrix. First column is the complete match, followed by one column for each capture group. For str_match_all, a list of character matrices.

Examples:
```{r str_match}


strings <- c(" 219 733 8965", "329-293-8753 ", "banana", "595 794 7569",
  "387 287 6718", "apple", "233.398.9187  ", "482 952 3315",
  "239 923 8115 and 842 566 4692", "Work: 579-499-7527", "$1000",
  "Home: 543.355.3679")
phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"

str_extract(strings, phone)

str_match(strings, phone)

# Extract/match all
str_extract_all(strings, phone)

str_match_all(strings, phone)
x <- c("<a> <b>", "<a> <>", "<a>", "", NA)
str_match(x, "<(.*?)> <(.*?)>")
str_match_all(x, "<(.*?)>")

str_extract(x, "<.*?>")

str_extract_all(x, "<.*?>")


```
## case ()

## invert_match ()

## modifiers ()

## stringr-data ()

## str_c ()

## str_conv ()

## str_count ()

## str_detect ()

## str_dup ()

## str_extract ()

## str_flatten ()

## str_glue ()

## str_length ()

## str_locate ()

## str_match ()

## str_order ()

## str_pad ()

## str_remove ()

## str_replace ()

## str_replace_na ()

## str_split ()

## str_sub ()

## str_subset ()

## str_trim ()

## str_trunc ()

## str_view ()

## str_wrap ()

## word ()
