# forcats {#forcats}

forcats is a non-core package in the tidyverse and all the main functions start with fct_ Although you can install forcats using install.packages("tidyverse"), it is loaded using library(forcats). 

## as_factor ()

as_factor (): Converts an input to a factor. as_factor() is a generic version of as.factor()

Description: Creates levels in the order in which they appear. This ensures means that as_factor(x) will always return the same result, regardless of the current locale.

Usage: 
```
as_factor(x, ...)
```
Example: 

```{r}
x <- c("e", "f", "g")
as_factor(x)
as.factor(x)
```

## fct_anonv ()


## fct_c ()
fct_c (): Concatenate factors, combining levels. 

Description: Allows the patching of factors having both different levels and are of different sources.


Usage: 
```
fct_c(...)
```

Example: 

```{r}
fr <- factor("r")
fs <- factor("s")
frs <- factor(c("r", "s"))

c(fr, fs, frs)
fct_c(fr, fs, frs)

```

## fct_collapse ()

Description: Collapse factor levels into manually defined groups.

Usage:
```{r}
wc_data <- read.csv2("https://raw.githubusercontent.com/fivethirtyeight/data/master/world-cup-predictions/wc-20140609-140000.csv", quote = "",header = TRUE, sep = ",")


wc_data <- mutate(wc_data, continent = fct_collapse(wc_data$country,
    "Asia" = c("Iran","Japan","South Korea","Russia"),
    "Africa" = c("Algeria", "Ivory Coast","Cameroon","Ghana","Nigeria"),
    "Australia" = c("Australia"),
    "Central America and the Antilles" = c("Costa Rica","Honduras"),
    "North America" = c("Mexico", "USA"),
    "South America" = c("Argentina","Brazil","Chile","Colombia","Ecuador"),
    "Europe" = c("Belgium", "Bosnia and Herzegovina","Croatia","England","Spain","France","Germany","Greece","Italy","Netherlands","Portugal","Switzerland","Uraguay" )
  ))
```

## fct_count ()
fct_count (): Count entries in a factor.

Description: Count entries in a factor.

Usage: 

```
fct_count(f, sort = FALSE)
```

Example: 

```{r}
f <- factor(sample(letters)[rpois(1000, 10)])
table(f)
fct_count(f)
fct_count(f, sort = TRUE)
```

## fct_drop ()
fct_drop (): ALlows you to drop unused levels.

Description: In comparison to base::droplevels(), does not drop NA levels that have values.

Usage: 

```
fct_drop(f, only)
```

Example: 

```{r}
f <- factor(c("a", "b"), levels = c("a", "b", "c"))
f
fct_drop(f)
```
```{r}
#Restrict which levels to drop
fct_drop(f, only = "a")
fct_drop(f, only = "c")
```
## fct_expand ()
fct_expand (): Add additional levels to a factor.

Description: Allows the expansion of levels to a chosen factor.

Usage: 

```
fct_expand(f, ...)
```

Example: 


```{r}
f <- factor(sample(letters[1:3], 20, replace = TRUE))
f
fct_expand(f, "d", "e", "f")
fct_expand(f, letters[1:6])
```

## fct_explicit_na ()
fct_explicit_na (): Make missing values explicit.

Description: Provides missing value an explicit factor level, ensuring that they appear in summaries and on plots.

Usage: 

```
fct_explicit_na(f, na_level = "(Missing)")

```

Example: 

```{r}
f1 <- factor(c("a", "a", NA, NA, "a", "a", NA, NA, "c", "a", "c", "b"))
table(f1)
```

```{r}
f2 <- fct_explicit_na(f1)
table(f2)
```



## fct_inorder ()
fct_inorder (): Reorder factors levels by first appearance or frequency.

Description: Reorders the factor levels by first appearance or frequency.

Usage: 

```
fct_inorder(f, ordered = NA)

fct_infreq(f, ordered = NA)
```

Example: 


```{r}
f <- factor(c("b", "b", "a", "c", "c", "c"))
f
fct_inorder(f)
fct_infreq(f)
```

```{r}
fct_inorder(f, ordered = TRUE)
```


## fct_lump ()

## fct_other ()
fct_other (): Replaces levels with "other."

Description: Makes it easier to convert selected levels to “other”

Usage: 

```
fct_other(f, keep, drop, other_level = "Other")

```

Example: 


```{r}
x <- factor(rep(LETTERS[1:9], times = c(40, 10, 5, 27, 1, 1, 1, 1, 1)))
fct_other(x, keep = c("A", "B"))
```

```{r}
fct_other(x, drop = c("A", "B", "C", "D"))
```

## fct_recode ()
fct_recode (): Change factor levels by hand.

Description: Allows the change of factor levels by hand.

Usage: 

```
fct_recode(.f, ...)
```

Example: 


```{r}
x <- factor(c("canada", "dog", "brazil", "cat"))
fct_recode(x, country = "canada", country = "brazil")
```

```{r}
# Keep in mind, if you name the level NULL it will be removed
fct_recode(x, NULL = "canada", country = "brazil")

```

## fct_relabel ()
fct_relabel(): ALlows the programmatic relabeling of levels.

Description: Automatically relabel factor levels, collapse as necessary.

Usage: 

```
fct_relabel(.f, .fun, ...)
```

Example: 

```{r}
x <- factor(letters[1:3])
x
```

```{r}
x %>% fct_relabel(function(x) paste0("-", x, "-"))
```

## fct_relevel ()
fct_relevel (): Reorder factor levels by hand.

Description: Allows you to move any number of levels to any location.

Usage: 

```
fct_relevel(.f, ..., after = 0L)

```

Example: 

```{r}
f <- factor(c("w", "x", "y", "z"))
fct_relevel(f)
```

```{r}
fct_relevel(f, "y")
```

```{r}
fct_relevel(f, "x", "w")
```

```{r}
# Move to the forth position
fct_relevel(f, "x", after = 3)
```

```{r}
# 'Inf' allows you to relevel to the end when there is an unknown number of levels
fct_relevel(f, "w", after = Inf)
```

## fct_reorder ()

## fct_rev ()

## fct_shift ()

## fct_shuffle ()

## fct_unify ()

## fct_unique ()

## gss_cat ()

## lvls ()

## lvls_union ()
