# lubridate {#lubridate}

The lubridate package makes working with date-time data easier. Base R commands for date-times are generally unintuitive not robust. According to Hadley Wickham, "lubridate has a consistent, memorable syntax, that makes working with dates fun instead of frustrating."  For additional information use help(package = lubridate) to bring up an overview of the package and its functions.

Load Packages:

```{r , message=FALSE, warning=FALSE}
library(lubridate)
```

Load data from the FiveThirtyEight article **"Some People Are Too Superstitious To Have A Baby On Friday The 13th"**

-  Since it is fairly clean data, I will create additional columns to showcase lubridate functionality below

-  I will use the now(), hour(), minute() and second() functions to add addditional metadata to the dataframe. Please see definitions below

-  Once the new columns are added, I will remove the original clean columns

```{r, message=FALSE, warning=FALSE}
US_births_2000 <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/births/US_births_1994-2003_CDC_NCHS.csv")%>%
                filter(year == 2000)%>%
                mutate(year = as.character(year),
                       month = as.character(month),
                       date_of_month = as.character(date_of_month)) %>%
               mutate(ymd.raw = paste(year, month, date_of_month, sep = '-'))%>%
               mutate(mdy.raw = paste(month,date_of_month,year, sep = '-'))%>%
               mutate(dmy.raw = paste(date_of_month,month, year,sep = '-'))%>%
              mutate(ymd.hms.raw = paste(year,month,date_of_month, hour(now()), minute(now()), second(now()),sep = '-'))%>%
              select(-c(year, month, date_of_month))

kable(head(US_births_2000,10),  caption = "2000 Birth Data")

```

##  ymd()

parses in order - year, month, date. I will apply to the "ymd.raw" column and save the results as "ymd.clean"

```{r}
cleaned_date_df <- US_births_2000%>%
                  mutate(ymd.clean = ymd(ymd.raw),
                         ymd.type = class(ymd.clean))%>%
                  select(-c(ymd.raw))
kable(head(cleaned_date_df,10),  caption = "ymd() results")
```

##  mdy()

parses in order- month, date, year. I will apply to the "mdy.raw" column and save the results as "mdy.clean." The mdy() function will save the results as year-month-date. This is helpful when trying to normalize data. As you can see below, the "ymd.clean" ="mdy.clean"

```{r}
cleaned_date_df <- cleaned_date_df%>%
                  mutate(mdy.clean = mdy(mdy.raw))%>%
                  select(-c(mdy.raw))
kable(head(cleaned_date_df,10),  caption = "mdy() results")
```

##  dmy()

parses in order- day, month, year. The dmy() function will save the results as year-month-date. This is helpful when trying to normalize data. As you can see below, the "ymd.clean" = "dmy.clean"

```{r}
cleaned_date_df <- cleaned_date_df%>%
                  mutate(dmy.clean = dmy(dmy.raw))%>%
                  select(-c(dmy.raw))
kable(head(cleaned_date_df,10),  caption = "dmy() results")
```

##  ymd_hms()

parses in order- year, month, day _ hour, minute, second

```{r}
cleaned_date_df <- cleaned_date_df%>%
                  mutate(ymd.hms.clean = ymd_hms(ymd.hms.raw))%>%
                  select(-c(ymd.hms.raw))
kable(head(cleaned_date_df,10),  caption = "ymd_hms() results")
```
```{r}
cleaned_date_df <- cleaned_date_df%>%
                  mutate(wday.clean = wday(ymd.clean))%>%
                  select(-c(day_of_week))
kable(head(cleaned_date_df,10),  caption = "wday() results")
```

## Additional Lubridate Functions

### Extract Information From Dates

-  tody() : returns today's date

-  year(): returns the year

-  month(): returns the month

```{r}
Sys.getlocale()
this_day <- today()
this_day
```

### Date-Time Information

*these were used to create the base dataframe*

-  day(): returns the day

-  wday(): day of the week

-  now(): date-time of the exact moment 

-  hour(): hour of the exact moment

-  minute(): minute of the exact moment 

-  second(): second of the exact moment 





